---
title: "Semantic Distances"
author: "Anna Chinni"
format: pdf
editor: visual
---

#Setup

```{r}
rm(list = ls())

# List of packages
packages <- c("corrr", "ggplot2", "lme4", "lmerTest", "FactoMineR", "grid","png","cowplot","magick", "purrr", "combinat", "ggimage","dplyr","tidyverse","knitr", "readxl", "writexl", "tidytext", "forcats", "ggrepel", "stringr", "jsonlite")
```

```{r}

install_and_load_packages <- function(packages) {
  for(package in packages) {
    if(!require(package, character.only = TRUE)) {
      install.packages(package, dependencies = TRUE)
      library(package, character.only = TRUE)
    }
  }
}

# Call the function with the list of packages
install_and_load_packages(packages)

```

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Functions

We operationalized semantic discriminability using the metric semantic distance

Semantic distance: - assumed associations ratings between features and concepts are normally distributed with a mean equal to the mean association rating and variance that is largest when the association rating is closest to the center of the rating scale. - defined: in an assignment, the absolute difference in the probabilities of each assignment being chosen by a random individual. - delta_x = (x1 + x4) - (x2 + x3)

x1 = assignment Concept 1 with Feature 1 x2 = assignment Concept 1 with feature 2 x3 = assignment Concept 2 with feature 1 x4 = assignment Concept 2 with feature 2

If delta_x \> 0 then feature 1 will be assigned with concept 1 and feature 2 will be assigned with concept 2. If delta_x \< 0 then feature 1 will be assigned with concept 2 and feature 2 will be assigned with concept 1.

- delta_x = \|Prob(delta_x\>0) - Prob(delta_x \<0)\|

```{r}
#semantic distance function 
get_semantic_distance <- function(concept1, concept2) {
  # Helper function for calculating the standard deviation
  sig <- function(xi) {
    return(max(1.4 * xi * (1 - xi), 0.1))
  }
  
  # Extract values from input vectors
  x1 <- concept1[1]
  x2 <- concept1[2]
  x3 <- concept2[1]
  x4 <- concept2[2]
  
  # Calculate semantic distance
  num <- (x1 + x4) - (x2 + x3)
  denom <- sqrt(sig(x1)^2 + sig(x2)^2 + sig(x3)^2 + sig(x4)^2)
  prob_dx_gt_zero <- pnorm(num / denom)
  prob_dx_lt_zero <- 1 - prob_dx_gt_zero
  semantic_distance <- abs(prob_dx_gt_zero - prob_dx_lt_zero)
  
  # Global assignment (1 = outer vs. 0 = inner edges)
  assignDiff <- (x1 + x4) - (x2 + x3)
  if (assignDiff > 0) {
    assign <- 1
  } else if (assignDiff < 0) {
    assign <- 0
  } else {
    assign <- NA
  }
  
  return(list(semantic_distance = semantic_distance, assign = assign))
}
```

Function for calculating entropy of each concept's distribution, takes a matrix where each concept is a row and each texture is a column with values as the average association rating for each pair.

```{r}
#entropy for each distribution
get_ents<-function(x){
  l={}
  for(item in x){
    l<-rbind(l,entropies_c[item])
    
    return(l)
  }
  
}

```

Function for Total Variation - difference in association distributions

```{r}
#TV
total_variation <- function(p_j1, p_j2) {
  0.5 * sum(abs(p_j1 - p_j2))
}
```

## Load Data

```{r}
current_dir<- getwd()

#data_dir <-paste0(current_dir,'/data/')

#data_files <-list.files(path = data_dir, full.names = TRUE, pattern="*.csv")
#print(paste0('we have data from ',length(data_files),' participants'))

#average association ratings
assoc <- read_csv("../data/tactile_associations_wide.csv")

#load in embedding space and key
#texture_space <- read_csv("triplet_textures_groupEmb_3d.csv")
#levels <- read_csv("d_levels.csv")

#textures
#textures <- read_csv("texture_list.csv")
textures <- data.frame(
  texture = colnames(assoc)[colnames(assoc) != "concept"]
)

```

## Calc entropies

```{r}
#define matrix structured tibble to plug into function for entropy values
ratings_mat <- assoc %>% 
  ungroup()

p_mat <- ratings_mat

#make sure that concepts column in character value

p_mat$concept<-as.character(p_mat$concept)

#check
class(p_mat$concept)

#make sure is matrix
#p_mat <- as.matrix(p_mat)

```

```{r}
#loop for normalizing ratings values
for(i in 1:nrow(p_mat)){
  p_mat[i,2:ncol(p_mat)]<- p_mat[i,2:ncol(p_mat)]/sum(p_mat[i,2:ncol(p_mat)])
}

#convert to base r data frame to work with original loop
p_mat <- as.data.frame(p_mat)

#empty vector to store entropy values
entropies = c()

#loop calculating entropy of each distribution --- KM original code for base r data frame
for (i in 1:nrow(p_mat)){
  logs = c()
  for(j in 2:length(p_mat[i,1:ncol(p_mat)])){
    if(p_mat[i,j]==0){
      logs<- append(logs,0)
    }else{
    logs<- append(logs, log(p_mat[i,j]))
    }
  }
  entropies<- append(entropies, sum(-logs*p_mat[i,2:ncol(p_mat)]))
}


#add names to entropy vector
names(entropies)<- ratings_mat$concept


#turn into tibble to merge with main associations ratings tibble
entropies_df <- tibble(
  concept = names(entropies),
  entropy = as.numeric(entropies)  
)

```

## Perceptual Distance

Perceptual distance - operationalized as the euclidean distance between visual textures in the raw embedding space estimated with Salmon in our triplet study.

```{r}
#Generate all texture pairs
texture_pairs <- combn(textures$texture, 2, simplify = FALSE)
```

```{r}
#perceptual distance for each texture pair

#clean texture names
levels <- levels %>% 
  mutate(
    texture =  gsub(".*(D\\d+).png", "\\1", levels$x)
  )
levels <- levels %>% 
  select(texture)

#remove extra row from embedding
texture_space <- texture_space %>% 
  slice(-1)

#combine with levels
texture_space <- bind_cols(levels, texture_space)

texture_space <- texture_space %>% 
  select(texture, `0`, `1`, `2`)

#rename
texture_space <- texture_space %>% 
  rename(
    x = `0`,
    y = `1`,
    z = `2`
  )
  
#use euclidean distance to calculate perceptual distance measure for each texture pair ----

#euclidean distance for 2 points in 3 dimensions
euclidean_distance <- function(row1, row2) {
  sqrt((row2$x - row1$x)^2 + (row2$y - row1$y)^2 + (row2$z - row1$z)^2)
}

#place to store the results
percept_distances_df <- data.frame(texture1 = character(0),
                                   texture2 = character(0),
                                   perceptual_distance = numeric(0))

# Loop through the texture pairs and calculate the distance
for (i in seq_along(texture_pairs)) {
  # Extract the texture names for this pair
  texture1 <- texture_pairs[[i]][1]
  texture2 <- texture_pairs[[i]][2]
  
  # Extract the rows corresponding to the textures
  row1 <- texture_space[texture_space$texture == texture1, ]
  row2 <- texture_space[texture_space$texture == texture2, ]
  
  #calc euclidean distance
  percept_distance <- euclidean_distance(row1, row2)
  
  #store
  percept_distances_df <- rbind(percept_distances_df, data.frame(texture1 = texture1, 
                                                                 texture2 = texture2, 
                                                                 perceptual_distance = percept_distance))
}

```

# Calc Semantic Distances

```{r}
#fruits
# fruits <- ratings_trials_exp %>% 
#   filter(concept %in% c("blueberry", "lemon", "mango", "strawberry", "watermelon", "apple", "banana", "cherry", "grape", "peach")) %>% 
#   group_by(concept, texture) %>% 
#   summarize(AverageRatings = mean(response), .groups = "drop")

# Generate all concept pairs
concepts <- unique(assoc$concept)
concept_pairs <- combn(concepts, 2, simplify = FALSE) 

# fruits_reshaped <- fruits %>%
#   spread(key = concept, value = AverageRatings)

# Initialize a dataframe to store results
semantic_distances_df <- data.frame(
  concept1 = character(),
  concept2 = character(),
  texture1 = character(),
  texture2 = character(),
  semantic_distance = numeric(),
  stringsAsFactors = FALSE
)

# Create an empty list to collect results
results_list <- vector("list", length = length(concept_pairs) * length(texture_pairs))
counter <- 1  # index for results_list

# Loop through each concept pair
for (pair in concept_pairs){
  # Loop through each combination of two textures
  for (textures in texture_pairs) {
    texture1 <- as.character(textures[[1]])
    texture2 <- as.character(textures[[2]])

    if (texture1 != texture2) {
      concept1_texture1_rating <- as.numeric(assoc[assoc$concept == pair[1], texture1])
      concept1_texture2_rating <- as.numeric(assoc[assoc$concept == pair[1], texture2])
      concept2_texture1_rating <- as.numeric(assoc[assoc$concept == pair[2], texture1])
      concept2_texture2_rating <- as.numeric(assoc[assoc$concept == pair[2], texture2])

      if (all(!is.na(c(concept1_texture1_rating, concept2_texture2_rating, 
                       concept2_texture1_rating, concept1_texture2_rating)))) {

        result <- get_semantic_distance(
                      c(concept1_texture1_rating, concept1_texture2_rating),
                      c(concept2_texture1_rating, concept2_texture2_rating)
        )

        # Store result as a list element
        results_list[[counter]] <- data.frame(
          concept1 = pair[1],
          concept2 = pair[2],
          texture1 = texture1,
          texture2 = texture2,
          semantic_distance = result$semantic_distance,
          stringsAsFactors = FALSE
        )
        counter <- counter + 1
      }
    }
  }
}

# Combine all results into a single data frame
semantic_distances_df <- do.call(rbind, results_list[1:(counter - 1)])

# Add identifier column with all the info
semantic_distances_df <- semantic_distances_df %>%
  mutate(identifier = paste(concept1, concept2, texture1, texture2, sep = "_"))

#add group name column
# semantic_distances_fruits_df <- semantic_distances_fruits_df %>% 
#   mutate(
#     group = "fruits"
#   )

##############

#max capacity for semantic discriminability for each concept pair
semantic_distances_df <- semantic_distances_df %>% 
  group_by(concept1, concept2) %>% 
  mutate(
    max_capacity = max(semantic_distance, na.rm = TRUE)
  ) %>% 
  ungroup()


#mean entropy

#add entropy values for each concept to df

#join to get entropy for concept_1
semantic_distances_df <- semantic_distances_df %>%
  left_join(entropies_df, by = c("concept1" = "concept")) %>%
  rename(entropy_c1 = entropy)

#join to get entropy for concept_2
semantic_distances_df <- semantic_distances_df %>%
  left_join(entropies_df, by = c("concept2" = "concept")) %>%
  rename(entropy_c2 = entropy)

#mean specificity
semantic_distances_df <- semantic_distances_df %>% 
  rowwise() %>% 
  mutate(
    mean_entropy = mean(c(entropy_c1, entropy_c2), na.rm = TRUE)
  ) %>% 
  ungroup()

#TV

#calc TV for every combination of concept pairs
#use normalized associations stored in p_mat - subset
# fruits_p_mat <- p_mat %>% 
#   filter(concept %in% c("blueberry", "lemon", "mango", "strawberry", "watermelon", "apple", "banana", "cherry", "grape", "peach"))

#calc TV

# Convert and reassign row names
p_mat_fixed <- as.matrix(p_mat[, -1])
rownames(p_mat_fixed) <- p_mat$concept

#place to store TVs
tv_results <- list()

# Loop through concept pairs
for (pair in concept_pairs) {
  concept1 <- pair[1]
  concept2 <- pair[2]
  
  # Get the rows from the matrix
  p_j1 <- p_mat_fixed[concept1, ]
  p_j2 <- p_mat_fixed[concept2, ]
  
  # Compute total variation
  tv <- total_variation(p_j1, p_j2)
  
  # Store the result as a list
  tv_results[[length(tv_results) + 1]] <- data.frame(
    concept1 = concept1,
    concept2 = concept2,
    TV = tv
  )
}

#combine into a data frame
tv_df <- do.call(rbind, tv_results)


#add column of normalized TV within fruits if want for seperate model
#fruits_tv_df$group_nTV <- fruits_tv_df$TV / sum(fruits_tv_df$TV)

tv_df$nTV <- (tv_df$TV - min(tv_df$TV)) / (max(tv_df$TV) - min(tv_df$TV))

#add TVs to main data frame
semantic_distances_df <- semantic_distances_df %>%
  left_join(tv_df, by = c("concept1", "concept2"))


#normalize entropy across pairs
#semantic_distances_fruits_df$group_n_mean_entropy <-  semantic_distances_fruits_df$mean_entropy/sum(semantic_distances_fruits_df$mean_entropy)

#subtract from 1 for specificity (so larger values are higher specificity)
#semantic_distances_fruits_df$group_specificity <- 1 - semantic_distances_fruits_df$group_n_mean_entropy

#add concept pair identifiers
semantic_distances_df$concept_pair <- paste(semantic_distances_df$concept1, semantic_distances_df$concept2, sep = "_")

```

## Combine Perceptual distances with Semantic Distances

```{r}
#add perceptual distances to each semantic distance df
semantic_distances_df <- merge(semantic_distances_df, percept_distances_df, by = c("texture1", "texture2"), all.x = FALSE, all.y = TRUE)

```

```{r}
#normalize mean entropy to it ranges from 0 to 1
semantic_distances_df$n_mean_entropy <- (semantic_distances_df$mean_entropy - min(semantic_distances_df$mean_entropy)) / (max(semantic_distances_df$mean_entropy) - min(semantic_distances_df$mean_entropy)) 
```

```{r}
#subtract from 1 for specificity (so larger values are higher specificity)
semantic_distances_df$specificity <- 1 - semantic_distances_df$n_mean_entropy
```

```{r}
#save semantic distance data frames as separate csv's

write.csv(semantic_distances_df, "../results/SemanticDistances_all.csv", row.names = FALSE)

```
