---
title: "VTBestRanges"
author: "Anna Chinni"
format: pdf
editor: visual
---

#Setup

```{r}
#rm(list = ls())

# List of packages
packages <- c("corrr", "ggplot2", "lme4", "lmerTest", "FactoMineR", "grid","png","cowplot","magick", "purrr", "combinat", "ggimage","dplyr","tidyverse","knitr", "readxl", "writexl", "tidytext", "forcats", "ggrepel", "stringr", "jsonlite")
```

```{r}

install_and_load_packages <- function(packages) {
  for(package in packages) {
    if(!require(package, character.only = TRUE)) {
      install.packages(package, dependencies = TRUE)
      library(package, character.only = TRUE)
    }
  }
}

# Call the function with the list of packages
install_and_load_packages(packages)

```

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

# Load in data

```{r}
data <- read_csv("../results/SemanticDistances_all.csv")
assoc <- read_csv("../data/tactile_associations_wide.csv")
```


# Select pairs that make sense given concept groups

Concepts groups

0 = clothes = c("dress", "pants", "shirt", "socks", "shoes")

1 = fruit1 = c("blueberry", "lemon", "mango", "strawberry", "watermelon")

2 = scenes = c("beach", "field", "ocean", "sky", "sunset")

3 = emotions = c("angry", "disgust", "fearful", "happy", "sad")

4 = locations = c("above", "below", "beside", "near", "far")

5 = times = c("dawn", "day", "dusk", "night", "noon")

6 = animals = c("bear", "bird", "lion", "frog", "fish")

7 = transit = c("airplane", "car", "boat", "truck", "train")

8 = fruit2 = c("apple", "banana", "cherry", "grape", "peach")

9 = vegetables = c("carrot", "celery", "corn", "eggplant", "mushroom")

10 = weather = c("blizzard", "drought", "hurricane", "lightning", "sandstorm")

11 = value = c("evil", "greed", "justice", "love", "peace")

12 = activity = c("driving", "eating", "sleeping", "leisure", "working")

13 = qualities = c("comfort", "efficiency", "speed", "safety", "reliability")

```{r}
clothes <- c("dress", "pants", "shirt", "socks", "shoes")
fruits <- c("blueberry", "lemon", "mango", "strawberry", "watermelon", "apple", "banana", "cherry", "grape", "peach")
scenes <- c("beach", "field", "ocean", "sky", "sunset")
emotions <- c("angry", "disgust", "fearful", "happy", "sad")
locations <- c("above", "below", "beside", "near", "far")
times <- c("dawn", "day", "dusk", "night", "noon")
animals <- c("bear", "bird", "lion", "frog", "fish")
transit <- c("airplane", "car", "boat", "truck", "train")
vegetables <- c("carrot", "celery", "corn", "eggplant", "mushroom")
weather <- c("blizzard", "drought", "hurricane", "lightning", "sandstorm")
values <- c("evil", "greed", "justice", "love", "peace")
activity <- c("driving", "eating", "sleeping", "leisure", "working")
qualities <- c("comfort", "efficiency", "speed", "safety", "reliability")
fruits_veggies <- c("blueberry", "lemon", "mango", "strawberry", "watermelon", "apple", "banana", "cherry", "grape", "peach", "carrot", "celery", "corn", "eggplant", "mushroom")
```


## Pairs

```{r}
#clothes
clothes <- combn(clothes, 2)
clothes <- as.data.frame(t(clothes))
colnames(clothes) <- c("concept1", "concept2")
clothes <- clothes %>% 
  mutate(
    category = "clothes"
  )

#fruits
fruits <- combn(clothes, 2)
fruits <- as.data.frame(t(fruits))
colnames(fruits) <- c("concept1", "concept2")
fruits <- fruits %>% 
  mutate(
    category = "fruits"
  )

#scenes
scenes <- combn(scenes, 2)
scenes <- as.data.frame(t(scenes))
colnames(scenes) <- c("concept1", "concept2")
scenes <- scenes %>% 
  mutate(
    category = "scenes"
  )

#emotions
emotions <- combn(emotions, 2)
emotions <- as.data.frame(t(emotions))
colnames(emotions) <- c("concept1", "concept2")
emotions <- emotions %>% 
  mutate(
    category = "emotions"
  )

#locations
locations <- combn(locations, 2)
locations <- as.data.frame(t(locations))
colnames(locations) <- c("concept1", "concept2")
locations <- locations %>% 
  mutate(
    category = "locations"
  )

#times
times <- combn(times, 2)
times <- as.data.frame(t(times))
colnames(times) <- c("concept1", "concept2")
times <- times %>% 
  mutate(
    category = "times"
  )

#animals
animals <- combn(animals, 2)
animals <- as.data.frame(t(animals))
colnames(animals) <- c("concept1", "concept2")
animals <- animals %>% 
  mutate(
    category = "animals"
  )

#transit
transit <- combn(transit, 2)
transit <- as.data.frame(t(transit))
colnames(transit) <- c("concept1", "concept2")
transit <- transit %>% 
  mutate(
    category = "transit"
  )

#vegetables
vegetables <- combn(vegetables, 2)
vegetables <- as.data.frame(t(vegetables))
colnames(vegetables) <- c("concept1", "concept2")
vegetables <- vegetables %>% 
  mutate(
    category = "vegetables"
  )

#weather
weather <- combn(weather, 2)
weather <- as.data.frame(t(weather))
colnames(weather) <- c("concept1", "concept2")
weather <- weather %>% 
  mutate(
    category = "weather"
  )

#values
values <- combn(values, 2)
values <- as.data.frame(t(values))
colnames(values) <- c("concept1", "concept2")
values <- values %>% 
  mutate(
    category = "values"
  )

#activity
activity <- combn(activity, 2)
activity <- as.data.frame(t(activity))
colnames(activity) <- c("concept1", "concept2")
activity <- activity %>% 
  mutate(
    category = "activity"
  )

#qualities
qualities <- combn(qualities, 2)
qualities <- as.data.frame(t(qualities))
colnames(qualities) <- c("concept1", "concept2")
qualities <- qualities %>% 
  mutate(
    category = "qualities"
  )

#fruits and veggies
fruits_veggies <- combn(fruits_veggies, 2)
fruits_veggies <- as.data.frame(t(fruits_veggies))
colnames(fruits_veggies) <- c("concept1", "concept2")
fruits_veggies <- fruits_veggies %>% 
  mutate(
    category = "fruits_veggies"
  )

```

```{r}
#combine
pairs <- bind_rows(clothes, fruits, scenes, emotions, locations, times, animals, transit, vegetables, weather, values, activity, qualities)

#fruits and veggies combined
#pairs <- bind_rows(clothes, fruits_veggies, scenes, emotions, locations, times, animals, transit, weather, values, activity, qualities)
```

```{r}
#add concept_pair key sorted alphabetically to match with data 
pairs <- pairs %>%
  mutate(
    concept_pair = if_else(concept1 < concept2,
                       paste(concept1, concept2, sep = "_"),
                       paste(concept2, concept1, sep = "_"))
  )
```

## Filter semantic distances to only have the within concept group pairs

```{r}
data <- data %>%
  semi_join(pairs, by = "concept_pair")
```

# Look at ranges for semantic distance

Only concept pairs with a max SemDist above .8 and a min below .2

```{r}
data <- data %>% 
  group_by(concept1, concept2) %>% 
  mutate(
    min_semd = min(semantic_distance)
  )
```

```{r}
data <- data %>% 
  filter(min_semd < .2)

data <- data %>% 
  filter(max_capacity > .65)
```

```{r}
data <- data %>%
  group_by(concept1, concept2) %>%
  mutate(range_semantic_distance = max(semantic_distance) - min(semantic_distance)) %>%
  ungroup()
```

```{r}
data <- data %>% 
  filter(range_semantic_distance >= .65)
```

```{r}
pair_ranges <- data %>%
  group_by(concept1, concept2) %>%
  summarise(
    range_semantic_distance = max(semantic_distance, na.rm = TRUE) -
                              min(semantic_distance, na.rm = TRUE),
    min_semd = min(semantic_distance),
    max_semd = max(semantic_distance),
    .groups = "drop"
  )
```

```{r}
unique_pairs <- data %>%
  distinct(concept1, concept2)

unique_pairs
```

```{r}
#write csv
#write_csv(data, "MC_results/best_ranges.csv")
```

## Find set of 6 textures that work

```{r}
# function to get MSE vs ideal uniform
get_mse <- function(distances) {
  ideal <- seq(0, 1, length.out = length(distances))
  mean((sort(distances) - ideal)^2)
}
```

```{r}
evaluate_set <- function(textures, df, min_range = .7, max_mse = .01) {
  sample_textures <- sample(textures, 6)
  pairs <- t(combn(sample_textures, 2)) |> as.data.frame()
  colnames(pairs) <- c("texture1", "texture2")
  
  #nake sure texture pairs match whats in sem dist data 
  joined <- pairs %>%
    left_join(df, by = c("texture1", "texture2")) %>%
    bind_rows(
      pairs %>%
        left_join(df, by = c("texture2" = "texture1", "texture1" = "texture2"))
    ) %>%
    filter(!is.na(semantic_distance)) %>%
    distinct(texture1, texture2, .keep_all = TRUE)
  
  if (nrow(joined) < 15) return(NULL)
  
  mse <- get_mse(joined$semantic_distance)
  range_val <- diff(range(joined$semantic_distance))
  
  #constraints
  if (range_val < min_range || mse > max_mse) return(NULL)
  
  list(set = sample_textures, mse = mse, range = range_val)
}

```

```{r}
find_best_set <- function(df, concept1, concept2, min_range = .7, max_mse = .01, n_tries = 5000) {
  subset_df <- df %>%
    filter(concept1 == !!concept1, concept2 == !!concept2)
  
  textures <- unique(c(subset_df$texture1, subset_df$texture2))
  best <- NULL
  
  for (i in seq_len(n_tries)) {
    res <- evaluate_set(textures, subset_df, min_range = min_range, max_mse = max_mse)
    
    if (!is.null(res)) {
      if (is.null(best) || res$mse < best$mse) best <- res
      #if (res$mse < threshold) break
    }
  }
  
  if (is.null(best)) return(NULL)
  tibble(
    concept1 = concept1,
    concept2 = concept2,
    textures = paste(best$set, collapse = ", "),
    mse = best$mse,
    range = best$range
  )
}

```

```{r}
#texture list
textures <- unique(c(data$texture1, data$texture2))
```

```{r}
#ideal set
ideal <- data.frame(
  rank = 1:15,
  semantic_distance = seq(0, 1, length.out = 15)
)
```

#Pairs

##Concrete 

Frog - Lion

Corn - Eggplant


##Abstract 

Evil - Peace

Comfort - Speed


# Frog Lion

```{r}
#select concept pair
frog_lion <- data %>%
  filter(
    (concept1 == "frog" & concept2 == "lion") |
    (concept1 == "lion" & concept2 == "frog")
  )
```

```{r}
frog_lion_results <- map_dfr(
  seq_len(nrow(frog_lion)),
  \(i) {
    row <- frog_lion[i, ]
    find_best_set(frog_lion, row$concept1, row$concept2, min_range = 0.71, max_mse = 0.02, n_tries = 2000)
  }
)
```

```{r}
#write csv
write_csv(frog_lion_results, "../results/frog_lion_results.csv")
```

```{r}
frog_lion_results <- read_csv("../results/frog_lion_results.csv")
```

```{r}
frog_lion_best <- frog_lion_results %>% 
  filter(range == max(range)) %>% 
  slice_min(order_by = mse, n = 1)
```

```{r}
ggplot(frog_lion_results, aes(x = range, y = mse, label = paste(concept1, concept2))) +
  geom_point(size = 3, color = "steelblue") +
  #geom_text(vjust = -0.5, size = 3) +
    geom_point(
    data = frog_lion_best,
    aes(x = range, y = mse),
    color = "red",
    size = 4
  ) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(limits=c(.0001, .008), expand = c(0,0)) +
  scale_x_continuous(limits=c(.5, 1), expand = c(0,0)) +
  labs(
    x = "Range of semantic distances (max - min)",
    y = "MSE vs. ideal uniform distribution",
    title = "Uniformity of semantic distances across 8-texture sets"
  )+
   theme(
    text = element_text(color = "black", size = 24),
    panel.background = element_blank(),       # Remove panel background
    plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    axis.line = element_line(color = "black") # Add axis lines
  ) 
```

```{r}
# Extract best textures
frog_lion_best_textures <- strsplit(frog_lion_best$textures, ",\\s*")[[1]]

# Subset your main data for those textures and concept pair
frog_lion_best_pairs <- frog_lion %>%
  filter(concept1 == frog_lion_best$concept1,
         concept2 == frog_lion_best$concept2,
         texture1 %in% frog_lion_best_textures,
         texture2 %in% frog_lion_best_textures) %>%
  distinct(texture1, texture2, .keep_all = TRUE)
```

```{r}
frog_lion_best_textures
```

```{r}
ggplot() +
  #geom_hline(yintercept = 0, color = "black") +
  # semdist points
  geom_point(
    data = frog_lion_best_pairs,
    aes(x = semantic_distance, y = .00005),
    size = 4,
    alpha = 0.6,
    color = "steelblue"
  ) +
  
  # ideal vertical dashes
  geom_point(
    data = ideal,
    aes(x = semantic_distance, y = .00005),
    shape = 124,   # vertical line shape in ggplot
    size = 6,      # controls dash length
    color = "black"
  ) +
  geom_text(
  data = ideal,
  aes(
    x = semantic_distance,
    y = 0,
    label = round(semantic_distance, 2)
  ),
  vjust = 1.8,        # pushes text below the line
  size = 6,
  color = "black"
) + 
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, .005), expand = c(0,0)) +
  theme_minimal() +
  labs(
    x = "",
    y = "",
    title = "Semantic Distance (0–1)"
  ) +
  theme(
    text = element_text(color = "black", size = 24),
    #panel.background = element_blank(),       # Remove panel background
    #plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    #panel.grid.minor = element_blank(),       # Remove minor grid lines
    # remove y axis entirely
  axis.title.y = element_blank(),
  axis.text.y  = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.y  = element_blank(),
  
  # remove x axis label (keep ticks if you want)
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_line(color = "black"),
  
  # background cleanup
  panel.background = element_blank(),
  plot.background  = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
    
    #axis.line = element_line(color = "black") # Add axis lines
  ) +
  coord_cartesian(clip = "off")

```

# Corn Eggplant

```{r}
#select concept pair
corn_eggplant <- data %>%
  filter(
    (concept1 == "corn" & concept2 == "eggplant") |
    (concept1 == "eggplant" & concept2 == "corn")
  )
```

```{r}
corn_eggplant_results <- map_dfr(
  seq_len(nrow(corn_eggplant)),
  \(i) {
    row <- corn_eggplant[i, ]
    find_best_set(corn_eggplant, row$concept1, row$concept2, min_range = 0.71, max_mse = 0.02, n_tries = 2000)
  }
)
```

```{r}
#write csv
write_csv(corn_eggplant_results, "../results/corn_eggplant_results.csv")
```

```{r}
corn_eggplant_results <- read_csv("../results/corn_eggplant_results.csv")
```

```{r}
corn_eggplant_best <- corn_eggplant_results %>% 
  filter(range == max(range)) %>% 
  slice_min(order_by = mse, n = 1)
```

```{r}
ggplot(corn_eggplant_results, aes(x = range, y = mse, label = paste(concept1, concept2))) +
  geom_point(size = 3, color = "steelblue") +
  #geom_text(vjust = -0.5, size = 3) +
  geom_point(
    data = corn_eggplant_best,
    aes(x = range, y = mse),
    color = "red",
    size = 4
  ) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(limits=c(.0001, .015), expand = c(0,0)) +
  scale_x_continuous(limits=c(.6, 1), expand = c(0,0)) +
  labs(
    x = "Range of semantic distances (max - min)",
    y = "MSE vs. ideal uniform distribution",
    title = "Uniformity of semantic distances across 8-texture sets"
  ) +
   theme(
    text = element_text(color = "black", size = 24),
    panel.background = element_blank(),       # Remove panel background
    plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    axis.line = element_line(color = "black") # Add axis lines
  ) 
```

```{r}
# Extract best textures
corn_eggplant_best_textures <- strsplit(corn_eggplant_best$textures, ",\\s*")[[1]]

# Subset your main data for those textures and concept pair
corn_eggplant_best_pairs <- corn_eggplant %>%
  filter(concept1 == corn_eggplant_best$concept1,
         concept2 == corn_eggplant_best$concept2,
         texture1 %in% corn_eggplant_best_textures,
         texture2 %in% corn_eggplant_best_textures) %>%
  distinct(texture1, texture2, .keep_all = TRUE)
```

```{r}
corn_eggplant_best_textures
```

```{r}
p_corn_eggplant_dist <- ggplot() +
  #geom_hline(yintercept = 0, color = "black") +
  # semdist points
  geom_point(
    data = corn_eggplant_best_pairs,
    aes(x = semantic_distance, y = .000075),
    size = .7,
    alpha = 0.6,
    color = "#1BA0C2"
  ) +
  
  # ideal vertical dashes
  geom_point(
    data = ideal,
    aes(x = semantic_distance, y = .000045),
    shape = 124,   # vertical line shape in ggplot
    size = 1,      # controls dash length
    color = "black"
  ) +
#   geom_text(
#   data = ideal,
#   aes(
#     x = semantic_distance,
#     y = 0,
#     label = round(semantic_distance, 2)
#   ),
#   vjust = 1.8,        # pushes text below the line
#   size = 6,
#   color = "black"
# ) + 
  geom_text(
  data = data.frame(x = c(0, 1), y = 0),
  aes(x = x, y = y, label = x),
  vjust = 1.8,
  size = 6,
  color = "black"
) +
  geom_text(
  data = data.frame(x = 0.5, y = 0),
  aes(x = x, y = y),
  label = "Semantic Discriminability",
  vjust = 3,   # push further below axis
  size = 6,
  color = "black"
)+
  scale_x_continuous(limits = c(0, 1), expand = c(0,0), breaks = seq(0, 1, by = 0.2)) +
  scale_y_continuous(limits = c(0, .005), expand = c(0,0)) +
  theme_minimal() +
  labs(
    x = "",
    y = "",
    title = "Semantic Distance (0–1)"
  ) +
  theme(
    plot.margin = margin(l = 20, r = 20, b=50),
    text = element_text(color = "black", size = 14),
    #panel.background = element_blank(),       # Remove panel background
    #plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    #panel.grid.minor = element_blank(),       # Remove minor grid lines
    # remove y axis entirely
  axis.title.y = element_blank(),
  axis.text.y  = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.y  = element_blank(),
  
  # remove x axis label (keep ticks if you want)
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_line(color = "black", linewidth = 0.2),
  
  # background cleanup
  panel.background = element_blank(),
  plot.background  = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
    
    #axis.line = element_line(color = "black") # Add axis lines
  ) +
  coord_cartesian(clip = "off")

p_corn_eggplant_dist
```

```{r}
ggsave(plot = p_corn_eggplant_dist,
       filename = "plots/p_corn_eggplant_dist.pdf",
       width = 2,
       height = 1,
       units = "in")
```

# Evil Peace

```{r}
#select concept pair
evil_peace <- data %>%
  filter(
    (concept1 == "evil" & concept2 == "peace") |
    (concept1 == "peace" & concept2 == "evil")
  )
```

```{r}
evil_peace_results <- map_dfr(
  seq_len(nrow(evil_peace)),
  \(i) {
    row <- evil_peace[i, ]
    find_best_set(evil_peace, row$concept1, row$concept2, min_range = 0.71, max_mse = 0.02, n_tries = 2000)
  }
)
```

```{r}
#write csv
write_csv(evil_peace_results, "../results/evil_peace_results.csv")
```

```{r}
evil_peace_results <- read_csv("../results/evil_peace_results.csv")
```

```{r}
evil_peace_best <- evil_peace_results %>% 
  filter(range == max(range))%>% 
  slice_min(order_by = mse, n = 1)
```

```{r}
ggplot(evil_peace_results, aes(x = range, y = mse, label = paste(concept1, concept2))) +
  geom_point(size = 3, color = "steelblue") +
  #geom_text(vjust = -0.5, size = 3) +
    geom_point(
    data = evil_peace_best,
    aes(x = range, y = mse),
    color = "red",
    size = 4
  ) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(limits=c(.0001, .015), expand = c(0,0)) +
  scale_x_continuous(limits=c(.6, 1), expand = c(0,0)) +
  labs(
    x = "Range of semantic distances (max - min)",
    y = "MSE vs. ideal uniform distribution",
    title = "Uniformity of semantic distances across 8-texture sets"
  )+
   theme(
    text = element_text(color = "black", size = 24),
    panel.background = element_blank(),       # Remove panel background
    plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    axis.line = element_line(color = "black") # Add axis lines
  ) 
```

```{r}
# Extract best textures
evil_peace_best_textures <- strsplit(evil_peace_best$textures, ",\\s*")[[1]]

# Subset your main data for those textures and concept pair
evil_peace_best_pairs <- evil_peace %>%
  filter(concept1 == evil_peace_best$concept1,
         concept2 == evil_peace_best$concept2,
         texture1 %in% evil_peace_best_textures,
         texture2 %in% evil_peace_best_textures) %>%
  distinct(texture1, texture2, .keep_all = TRUE)
```

IMPORTANT: G75 and S22 produce a tie for optimal assignment for the concept set evil and peace. Therefore we selected the next set with the highest range of semantic distances that did not include the G75 and S22 pairing. This was the 3rd best range with a minuscule difference and had a lower MSE. That is the set we selected for the experiment to avoid over complicating the optimal assignment for coding accuracy. 

```{r}
evil_peace_best_textures
```

```{r}
ggplot() +
  #geom_hline(yintercept = 0, color = "black") +
  # semdist points
  geom_point(
    data = evil_peace_best_pairs,
    aes(x = semantic_distance, y = .00005),
    size = 4,
    alpha = 0.6,
    color = "steelblue"
  ) +
  
  # ideal vertical dashes
  geom_point(
    data = ideal,
    aes(x = semantic_distance, y = .00005),
    shape = 124,   # vertical line shape in ggplot
    size = 6,      # controls dash length
    color = "black"
  ) +
  geom_text(
  data = ideal,
  aes(
    x = semantic_distance,
    y = 0,
    label = round(semantic_distance, 2)
  ),
  vjust = 1.8,        # pushes text below the line
  size = 6,
  color = "black"
) + 
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, .005), expand = c(0,0)) +
  theme_minimal() +
  labs(
    x = "",
    y = "",
    title = "Semantic Distance (0–1)"
  ) +
  theme(
    text = element_text(color = "black", size = 24),
    #panel.background = element_blank(),       # Remove panel background
    #plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    #panel.grid.minor = element_blank(),       # Remove minor grid lines
    # remove y axis entirely
  axis.title.y = element_blank(),
  axis.text.y  = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.y  = element_blank(),
  
  # remove x axis label (keep ticks if you want)
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_line(color = "black"),
  
  # background cleanup
  panel.background = element_blank(),
  plot.background  = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
    
    #axis.line = element_line(color = "black") # Add axis lines
  ) +
  coord_cartesian(clip = "off")

```

# Comfort Speed

```{r}
#select concept pair
comfort_speed <- data %>%
  filter(
    (concept1 == "comfort" & concept2 == "speed") |
    (concept1 == "speed" & concept2 == "comfort")
  )
```

```{r}
comfort_speed_results <- map_dfr(
  seq_len(nrow(comfort_speed)),
  \(i) {
    row <- comfort_speed[i, ]
    find_best_set(comfort_speed, row$concept1, row$concept2, min_range = 0.71, max_mse = 0.02, n_tries = 2000)
  }
)

```

```{r}
#write csv
write_csv(comfort_speed_results, "../results/comfort_speed_results.csv")
```

```{r}
comfort_speed_results <- read_csv("../results/comfort_speed_results.csv")
```

```{r}
comfort_speed_best <- comfort_speed_results %>% 
  filter(range == max(range)) %>% 
  slice_min(order_by = mse, n = 1)
```

```{r}
ggplot(comfort_speed_results, aes(x = range, y = mse, label = paste(concept1, concept2))) +
  geom_point(size = 3, color = "steelblue") +
  #geom_text(vjust = -0.5, size = 3) +
  geom_point(
    data = comfort_speed_best,
    aes(x = range, y = mse),
    color = "red",
    size = 4
  ) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(limits=c(.05, .005), expand = c(0,0)) +
  scale_x_continuous(limits=c(.65, .96), expand = c(0,0)) +
  labs(
    x = "Range of semantic distances (max - min)",
    y = "MSE vs. ideal uniform distribution",
    title = "Uniformity of semantic distances across 8-texture sets"
  )+
   theme(
    text = element_text(color = "black", size = 24),
    panel.background = element_blank(),       # Remove panel background
    plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    axis.line = element_line(color = "black") # Add axis lines
  ) 
```

```{r}
# Extract best textures
comfort_speed_best_textures <- strsplit(comfort_speed_best$textures, ",\\s*")[[1]]

# Subset your main data for those textures and concept pair
comfort_speed_best_pairs <- comfort_speed %>%
  filter(concept1 == comfort_speed_best$concept1,
         concept2 == comfort_speed_best$concept2,
         texture1 %in% comfort_speed_best_textures,
         texture2 %in% comfort_speed_best_textures) %>%
  distinct(texture1, texture2, .keep_all = TRUE)
```

```{r}
comfort_speed_best_textures
```

```{r}
ggplot() +
  #geom_hline(yintercept = 0, color = "black") +
  # semdist points
  geom_point(
    data = comfort_speed_best_pairs,
    aes(x = semantic_distance, y = .00005),
    size = 4,
    alpha = 0.6,
    color = "steelblue"
  ) +
  
  # ideal vertical dashes
  geom_point(
    data = ideal,
    aes(x = semantic_distance, y = .00005),
    shape = 124,   # vertical line shape in ggplot
    size = 6,      # controls dash length
    color = "black"
  ) +
  geom_text(
  data = ideal,
  aes(
    x = semantic_distance,
    y = 0,
    label = round(semantic_distance, 2)
  ),
  vjust = 1.8,        # pushes text below the line
  size = 6,
  color = "black"
) + 
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, .005), expand = c(0,0)) +
  theme_minimal() +
  labs(
    x = "",
    y = "",
    title = "Semantic Distance (0–1)"
  ) +
  theme(
    text = element_text(color = "black", size = 24),
    #panel.background = element_blank(),       # Remove panel background
    #plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    #panel.grid.minor = element_blank(),       # Remove minor grid lines
    # remove y axis entirely
  axis.title.y = element_blank(),
  axis.text.y  = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.y  = element_blank(),
  
  # remove x axis label (keep ticks if you want)
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_line(color = "black"),
  
  # background cleanup
  panel.background = element_blank(),
  plot.background  = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
    
    #axis.line = element_line(color = "black") # Add axis lines
  ) +
  coord_cartesian(clip = "off")

```



```{r}
best_sets <- bind_rows(frog_lion_best, corn_eggplant_best, evil_peace_best, comfort_speed_best)
```

```{r}
#write csv
write_csv(best_sets, "../results/exp2_sets.csv")
```

```{r}
best_sets <- read_csv("../results/exp2_sets.csv")
```

```{r}
ggplot(best_sets, aes(x = range, y = mse, label = paste(concept1, concept2))) +
  geom_point(size = 3, color = "steelblue") +
  geom_text(
    aes(label = paste(concept1, concept2)),
    vjust = -0.8,
    size = 6
  ) +
  #geom_text(vjust = -0.5, size = 3) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(limits=c(0, .05), expand = c(0,0)) +
  scale_x_continuous(limits=c(.65, 1), expand = c(0,0)) +
  labs(
    x = "Range of semantic distances",
    y = "MSE from ideal uniform distribution",
    title = "Uniformity of semantic distances across 8-texture sets"
  )+
   theme(
    text = element_text(color = "black", size = 24),
    panel.background = element_blank(),       # Remove panel background
    plot.background = element_blank(),        # Remove overall plot background
    #panel.grid.major = element_blank(),       # Remove major grid lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    axis.line = element_line(color = "black") # Add axis lines
  ) 
```

# Look at Associations

```{r}
#separate textures in to distinct columns
best_sets_assoc <- best_sets %>%
  separate(
    textures,
    into = paste0("texture", 1:6),
    sep = ",\\s*"
  )%>% 
  pivot_longer(
    cols = starts_with("texture"),
    names_to = "texture_pos",
    values_to = "texture"
  )
```

```{r}
assoc_long <- assoc %>%
  pivot_longer(
    cols = -concept,
    names_to = "texture",
    values_to = "association"
  )
```

```{r}
best_sets_assoc <- best_sets_assoc %>%
  mutate(pair = interaction(concept1, concept2, sep = "_")) %>%
    pivot_longer(
    cols = c(concept1, concept2),
    names_to = "concept_role",
    values_to = "concept"
  ) %>%
  left_join(
    assoc_long,
    by = c("concept", "texture")
  )

best_sets_assoc$texture <- as_factor(best_sets_assoc$texture)
```

```{r}
concept1_assoc <- best_sets_assoc %>%
  filter(concept_role == "concept1") %>%
  select(pair, texture, association) %>%
  rename(concept1_assoc = association)
```

```{r}
best_sets_assoc  <- best_sets_assoc %>%
  left_join(concept1_assoc, by = c("pair", "texture"))
```

Add image paths

```{r}
# best_sets_assoc <- best_sets_assoc %>%
#   group_by(pair) %>%
#   arrange(concept1_assoc, .by_group = TRUE) %>%
#   mutate(
#     texture = reorder_within(texture, -concept1_assoc, pair),
#     texture_for_image = texture  # <- this keeps the ordering for plotting
#   ) %>%
#   ungroup() %>%
#   mutate(
#     texture_id = sub("___.*$", "", texture_for_image),  # for filenames only
#     image_path = paste0("/Users/annachinni/Dropbox/TextureSemantics/VisualSemanticDist/NBT_53_bars/",
#                         texture_id, ".png"),
#     image_y = 0
#   )
```

```{r}
best_sets_assoc <- best_sets_assoc %>%
  group_by(pair) %>%
  arrange(concept1_assoc, .by_group = TRUE) %>%
  mutate(
    texture = reorder_within(
      texture,
      -concept1_assoc,
      pair
    )
  ) %>%
  ungroup()
```

```{r}
all_pairs <- unique(best_sets_assoc$pair)
```

```{r}
plots <- list()

for (p in all_pairs) {

  df_pair <- best_sets_assoc %>% 
    filter(pair == p) %>%
    group_by(concept_role) %>%
    mutate(
      texture = fct_reorder(texture, association, .desc = TRUE),
      #texture_for_image = texture,   
      #image_y = -0.1                    # images sit on baseline
    ) %>%
    ungroup()
    #mutate(
      #label_y = -0.235  # <- position for the x-axis labels
    #)

  plt <- ggplot(df_pair, aes(x = texture, y = association)) +
    ## ---- IMAGES ----
    # geom_image(
    #   aes(x = texture_for_image, y = image_y, image = image_path),
    #   size = 0.08,
    #   by = "width"
    # ) +
    # geom_text(
    #   aes(x = texture, y = label_y, label = sub("___.*$", "", texture)),
    #   inherit.aes = FALSE,
    #   angle = 0,
    #   hjust = 1,
    #   size = 5,
    #   color = NA
    # ) +
    geom_col(width = 0.6, fill = "steelblue") +
    #geom_hline(yintercept = 0, color = "black", linewidth = 0.6) +
    # geom_segment(aes(x = 0.1, xend = 0.1, y = 0, yend = 1), 
    #            color = "black", linewidth = 0.6) +
    facet_grid(~concept, scales = "free_x") +
    scale_x_discrete(
      labels = function(x) sub("___.*$", "", x),
      expand = c(.1, 0)
    ) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    coord_cartesian(clip = "off") +
    labs(
      x = "Texture",
      y = "Association strength",
      title = paste("Pair:", p)
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.margin = margin(t = 10, r = 20, b = 100, l = 20),
      axis.text.x = element_text(color = "black", size = 14, margin = margin(t = 40)),
      axis.text.y = element_text(color = "black", size = 14),
      axis.title = element_text(color = "black", size = 16, face = "bold"),
      strip.text = element_text(color = "black", size = 14, face = "bold"),
      plot.title = element_text(color = "black", size = 18, face = "bold"),
      panel.spacing = unit(2, "lines"),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      panel.grid = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "none",
      #axis.line.x = element_line(color = NA),
      #axis.line.y = element_line(color = NA)
    )

  plots[[p]] <- plt
}
```

```{r}
plots[["comfort_speed"]]
```

```{r}
# Make sure you have a directory to save the plots
output_dir <- "/Users/annachinni/Dropbox/TextureSemantics/VisualSemanticDist/Plots/"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Loop over your plots list and save each
for (p in names(plots)) {
  # Construct the filename
  file_name <- paste0(output_dir, p, "_assoc.pdf")
  
  # Save the plot
  ggsave(
    filename = file_name,
    plot = plots[[p]],
    width = 12,       # adjust as needed
    height = 6,       # adjust as needed
    dpi = 300
  )
}

```

#checking semantic distances

```{r}
bernoulli <- function(x){
  x * (1 - x)
}
```

```{r}
v <- seq(0, 1, length.out = 500)

plot(v, bernoulli(v), type = "l")
```

```{r}
sig <- function(xi) {
  pmax(1.4 * xi * (1 - xi), 0.1)  # pmax works element-wise
}
```

```{r}
plot(v, sig(v), type = "l")
```

```{r}
sig(.6)
```

sig is stand in for standard deviaion, bernoulli as approach 0 or 1 probability becomes more certain, so smaller sd/measure of noise in the center there is the most uncertainty its the noisiest.

then we are summing the associations (x1 + x4) - (x2 + x3) (essentially) so we need a variance sum

then calc variance from the individual standard deviations which is just sd\^2

you can sum variance normally

then we convert back to a standard deviation by square rooting our sum

this is a combined standard deviation for the combined association measures

then divide associaiton (signal) by sd (noise, same units as signal)

ratio of signal or measure to the noise -\> bigger value (difference, due to small noise) higher certainty, and vice versa

we then turn it into a probability like measure

```{r}
get_semantic_distance <- function(concept1, concept2) {
  # Helper function for calculating the standard deviation
  sig <- function(xi) {
    return(max(1.4 * xi * (1 - xi), 0.1))
  }
  
  # Extract values from input vectors
  x1 <- concept1[1]
  x2 <- concept1[2]
  x3 <- concept2[1]
  x4 <- concept2[2]
  
  # Calculate semantic distance
  num <- (x1 + x4) - (x2 + x3)
  denom <- sqrt(sig(x1)^2 + sig(x2)^2 + sig(x3)^2 + sig(x4)^2)
  prob_dx_gt_zero <- pnorm(num / denom)
  prob_dx_lt_zero <- 1 - prob_dx_gt_zero
  semantic_distance <- abs(prob_dx_gt_zero - prob_dx_lt_zero)
  
  # Global assignment (1 = outer vs. 0 = inner edges)
  assignDiff <- (x1 + x4) - (x2 + x3)
  if (assignDiff > 0) {
    assign <- 1
  } else if (assignDiff < 0) {
    assign <- 0
  } else {
    assign <- NA
  }
  
  return(list(semantic_distance = semantic_distance, assign = assign))
}
```
